Formal Modeling and Verification of a Rate-Monotonic Scheduling Implementation with Real-Time Maude

### 2015-11-06 RMS/submission.tex
polished until first part of Section.Modeling.

It is also crucial to ensure the reliability of the implementation of
RMS when it serves in a safety-critical system. When analyzing a
realistic implementation instead of the algorithm, theoretical results
may be no more applicable. For instance, switching overhead may lead
to unschedulability even if the task set satisfies the schedulability
condition derived from algorithm analysis. Extensive effort to apply
formal methods, such as model checking and theorem proving, has been
made to analyze safety-critical systems for the past few
years~\cite{DBLP:journals/tie/MiyawakiMSYV05,DBLP:journals/iandc/MeseguerR13,DBLP:journals/cacm/Leroy09,DBLP:conf/sosp/KleinEHACDEEKNSTW09}. However,
as far as we know, few~\cite{DBLP:conf/iceccs/CuiDT14,TianD2011}
attempt to analyze the RMS algorithm, while no work for
implementations of RMS is found.

@@

However in this paper, we consider an implementation instead of the
RMS algorithm itself, thus the model would be more complicated than
this standard, ideal setting. Assumption~(A1) will be modified because
of the interrupt masking mechanism, while (A2) and (A4) are relaxed to
obtain a more realistic analysis model.

@@

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given $t$, any of them
may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.

Real-Time Maude also supports specifications in \emph{object-oriented}
style.  A class declaration $\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_1$ to $val_n$ are the values of the attributes $att_1$ to
$att_n$. A \emph{subclass} inherits all the attributes and rules of
its superclasses.

\subsubsection{Formal Analysis}
Real-Time Maude provides many useful commands and tools to analyze a
model. For example, \verb|rewrite| allows to execute the model,
symbolically; given an initial state, \verb|search| is used to search
reachable states satisfying the desired properties; the Maude's
\emph{Inductive Theorem Prover} (ITP) can be applied to interactively
prove properties written in \emph{membership equational logic}.

In this paper, we only consider Real-Time Maude's \emph{linear
  temporal logic model checker}, which analyzes whether \emph{each}
behavior satisfies a temporal logic formula. \emph{State propositions}
are terms of sort \verb|Prop|. Their semantics is defined by equations
of the form $\texttt{ceq } statePattern \texttt{ |= } prop \texttt{ =
} b \texttt{ if } cond$, with $b$ a term of sort \verb|Bool|, stating
that $prop$ evaluates to $b$ in states which are instances of
$statePattern$ provided the condition $cond$ holds. These equations
together define $prop$ to hold in all states $t$ that make $t \texttt{
  |= } prop$ evaluate to \verb|true|. A temporal logic \emph{formula}
is constructed by state propositions and temporal logic operators such
as \verb|True|, \verb|False|, \verb|~|(negation), \verb|/\|,
\verb|\/|, \verb|->|(implication), \verb|[]|(``always''),
\verb|<>|(``eventually'') and \verb|U|(``until''). Real-Time Maude
supports both \emph{timed} and \emph{untimed LTL model checking}. The
untimed model checking command
\begin{alltt}
  (mc \(t\) |=u \(formula\) .)
\end{alltt}
checks whether the temporal logic $formula$ holds in all behaviors
starting from the initial state $t$, \emph{with no time limit}.

\section{The Implementation of RMS}
\label{s:imp}
The implementation is in a realistic avionic operating system.
Interrupts would be triggered by the clock every $T$, which we call
\emph{interrupt cycle}. When an interrupt request occurs, if the
system is interruptable, i.e. the interrupt mask is cleared, the
handler function $schedule()$ will be invoked; otherwise $schedule()$
will be pending until the interrupt mask becomes cleared.  The
implementation is shown as $schedule()$ in Figure~\ref{a:schedule},
where $taskList$ is the list of periodic tasks to be scheduled. We
assume that the list is descending order of priority, and both
variables $taskList$ and $timer$ are global. In this implementation,
there is only one kind of interrupt, the period $T_i$ of each task is
a multiple of $T$, and the tasks are independent, meeting
assumption~(A2).

\begin{figure}[!t]
\hrule\vspace{1mm}
  \begin{algorithmic}[1]
\Function{$schedule$}{$ $}{}
  \State \Call{$int\_o\!f\!\!f$}{$ $}; \Comment{to disable interrupts} \label{l:1stline}
  \State \Call{$updateStatus$}{$taskList$}; \label{l:updatestatus}
  \State $timer = timer + 1$; \label{l:timer} \label{l:inc}
  \State $p = taskList$;
  \While{$p$} \label{l:startrun1st}
    \If{$p\rightarrow status == \textit{INTERRUPT}$}
      \State \Return;
    \ElsIf{$p\rightarrow status == \textit{READY}$}      
      \State $p\rightarrow status = \textit{RUNNING}$;
      \State \Call{$int\_on$}{$ $}; \Comment{to enable interrupts} \label{l:endrun1st}
      \State $p\rightarrow function()$; \Comment{to execute the task} \label{l:function}
      \State \Call{$int\_o\!f\!\!f$}{$ $};
      \State $p\rightarrow status = \textit{DORMANT}$;
    \EndIf
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
\Function{$updateStatus$}{$p$}
  \While{$p$}
    \If{$p\rightarrow status == \textit{RUNNING}$} \label{l:startupdate}
      \State $p\rightarrow status = \textit{INTERRUPT}$;
    \EndIf
    \If{$timer~\%~(p\rightarrow period) == 0$} \Comment{the task should be initiated}
      \If{$p\rightarrow status == \textit{DORMANT}$} \Comment{the previous job finishes} 
        \State $p\rightarrow status = \textit{READY}$;
      \Else \Comment{the status is \textit{READY} or \textit{INTERRUPT}}
	\State \Call{$reportTaskError$}{$p$}; \Comment{the task misses its deadline}
      \EndIf
    \EndIf \label{l:endupdate}
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
  \end{algorithmic}
\hrule
  \caption{The C-Like Pseudocode of $schedule()$}
  \label{a:schedule}
\end{figure}

In Figure~\ref{a:schedule}, the handler function $scheduler()$
first updates status of all tasks in $taskList$ via function
$updateStatus()$. Then it traverses the list to execute the ready
tasks one by one, or to do a return when encountering an
interrupted\footnote{Note that the status \textit{INTERRUPT} indicates
  the task is interrupted for the moment, or was interrupted before
  but its execution is not complete yet.} task. The function
$updateStatus()$ updates each task in two steps: firstly, if the task
is running, it becomes interrupted; secondly for the task at its
initiation time, if its previous job is complete, it would be set to
be ready, otherwise it misses its deadline, producing an error. Notice
that $schedule()$ is invoked only when the periodic interrupt is
handled. Its execution cannot be interrupted when it is updating
status of tasks or finding the next task to execute, however, it can
be interrupted while executing some task (Line~\ref{l:function}). This
allows the execution of $schedule()$ to be nested.

In the rest of this paper, we use \emph{scheduling} to refer to the
stage, from the moment when a pending interrupt request is detected,
to the moment when the first should-be-run periodic task starts
executing, i.e. Line~\ref{l:function} in
Figure~\ref{a:schedule}. Therefore, \emph{scheduling time} mainly
consists of three parts: (i) the time for switching context from the
running task, possibly none, to $schedule()$ when an interrupt request
is handled, (ii) the time spent by $schedule()$ finding and setting
the \emph{first} should-be-run periodic task
(Lines~\ref{l:1stline}-\ref{l:endrun1st} in
Figure~\ref{a:schedule}), and (iii) the time for switching context
from $schedule()$ to that first should-be-run task. \emph{Switching}
refers to the stage, from the moment when a periodic task completes
its execution, to the moment when the next should-be-run periodic task
starts executing. \emph{Switching time} thus also consists of three
parts: (i) the time for switching context from the finished task back
to $schedule()$, (ii) the time spent by $schedule()$ finding and
setting the \emph{next} should-be-run periodic task, and (iii) the
time for switching context from $schedule()$ to that next
should-be-run task.

\section{Formal Modeling of the Implementation}
\label{s:formalism}
We model the implementation under the following assumptions:

(A1') Computation requirement $C_i$ for each task $\tau_i$ is not
constant. It varies randomly between minimum $C^{min}_i$ and maximum
$C^{max}_i$, i.e. within an interval $[C^{min}_i, C^{max}_i]$.

(A2) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly.

(A3') Scheduling time and switching time are considered, while other
overhead is ignored.

These assumptions make our model different from the standard one. For
instance, if an interrupt request occurs while switching is performed,
jobs corresponding to task $\tau_i$ may not be initiated at exactly
$kT_i$. They should wait until switching finishes and the interrupt
mask is cleared. Note that (A2) says that jobs are ready to run at
their initiation times, however, no one can start running at exactly
its initiation time, because scheduling takes time.

\jx{TODO. A figure is needed here to demonstrate the scheduling under
  the three assumptions above.}

>>>>

It is even more crucial to ensure the reliability of an implementation
instead of the algorithm, when RMS serves in a safety-critical
system. When analyzing a realistic implementation, theoretical results
may be no more applicable. For instance, even though the conditions
derived from algorithm analysis are satisfied, schedulability can be
broken by overhead in the system, or by the interrupt masking
mechanism which may delay interrupt handling. On the other hand,
correctness of the implementation with respect to the algorithm is
difficult to be verified by the traditional methods such as testing
and simulation, since they are incomplete. Extensive effort to apply
formal methods, such as model checking and theorem proving, has been
made to analyze safety-critical systems for the past few
years~\cite{DBLP:journals/tie/MiyawakiMSYV05,DBLP:journals/iandc/MeseguerR13,DBLP:journals/cacm/Leroy09,DBLP:conf/sosp/KleinEHACDEEKNSTW09}. However,
as far as we know, few~\cite{DBLP:conf/iceccs/CuiDT14,TianD2011}
attempt to analyze the RMS algorithm, while no work for
implementations of RMS is found.

@@

A simple example showing this algorithm is depicted in
Figure~\ref{sf:rmsalg}. However in this paper, we consider an
implementation instead of the RMS algorithm itself, thus the model
would be more complicated than this standard, ideal
setting. Assumption~(A1) will be modified because of the interrupt
masking mechanism, while (A2) and (A4) are relaxed to obtain a more
realistic analysis model.

@@

In rewriting logic, rewrite rules are applied non-deterministically,
that is, when several rules can be applied on a given $s$, any of them
may be chosen. Hence non-deterministic behaviors can be modeled
naturally in Real-Time Maude.  Real-Time Maude also supports
specifications in \emph{object-oriented} style.  A class declaration
$\texttt{class }C\texttt{ |
}att_1\texttt{:}s_1\texttt{,}\ldots\texttt{,}att_n\texttt{:}s_n$
defines a class $C$ with attributes $att_1$ to $att_n$ of sorts $s_1$
to $s_n$, respectively. An \emph{object} of class $C$ is represented
as a term $\texttt{< } O\texttt{:} C \texttt{ | }
att_1\texttt{:}val_1\texttt{,} \ldots
\texttt{,}att_n\texttt{:}val_n\texttt{ >}$ of sort \verb|Object|,
where $O$, of sort \verb|Oid|, is the object's \emph{identifier}, and
$val_i$ is the value of the attribute $att_i$ with $i\in [1,n]$. A
\emph{subclass} inherits all the attributes and rules of its
superclasses.

\subsubsection{Formal Analysis}
Real-Time Maude provides many useful commands and tools to analyze a
given model. For example, \verb|rewrite| allows to execute the model,
symbolically; given an initial state, \verb|search| is used to search
reachable states satisfying the desired properties; the Maude's
\emph{Inductive Theorem Prover} (ITP) can be applied to interactively
prove properties written in \emph{membership equational logic}.

In this paper, we only consider Real-Time Maude's \emph{Linear
  Temporal Logic (LTL) model checker}, which analyzes whether
\emph{each} behavior satisfies a temporal logic formula. \emph{State
  propositions} are defined as terms of sort \verb|Prop|. Their
semantics is defined by conditional equations of the form $\texttt{ceq
} statePattern \texttt{ |= } prop \texttt{ = } b \texttt{ if } cond$,
with $b$ a term of sort \verb|Bool|, stating that $prop$ evaluates to
$b$ in states which are instances of $statePattern$ provided the
condition $cond$ holds. These equations together define $prop$ to hold
in all states $s$ that make $s \texttt{ |= } prop$ evaluate to
\verb|true|. A temporal logic \emph{formula} is constructed by state
propositions and temporal logic operators such as \verb|True|,
\verb|False|, \verb|~|(negation), \verb|/\|, \verb|\/|,
\verb|->|(implication), \verb|[]|(``always'') and
\verb|U|(``until''). Real-Time Maude supports both \emph{timed} and
\emph{untimed LTL model checking}. The untimed model checking command
\begin{alltt}
  (mc \(s\) |=u \(formula\) .)
\end{alltt}
checks whether the temporal logic $formula$ holds in all behaviors
starting from the initial state $s$, \emph{with no time limit}.

\section{The Implementation of RMS}
\label{s:imp}
The implementation is in a realistic avionic control system.
Interrupts would be triggered by, and only by, the clock every $T$,
which we call \emph{interrupt cycle}. When an interrupt request
occurs, if the system is interruptable, i.e. the interrupt masking bit
is cleared, the handler function $schedule()$ will be invoked;
otherwise $schedule()$ will be pending until the interrupt masking bit
becomes cleared.  The implementation is shown as $schedule()$ in
Figure~\ref{a:schedule}, where $taskList$ is the list of periodic
tasks to be scheduled. We assume that the list is in descending order
of priority, and both variables $taskList$ and $timer$ are global. In
this implementation, there is only one kind of interrupt, the period
$T_i$ of each task is a multiple of $T$, and the tasks are
independent, meeting assumption~(A3).

\begin{figure}[!t]
\hrule\vspace{1mm}
  \begin{algorithmic}[1]
\Function{$schedule$}{$ $}{}
  \State \Call{$int\_o\!f\!\!f$}{$ $}; \Comment{to disable interrupts} \label{l:1stline}
  \State \Call{$updateStatus$}{$taskList$}; \label{l:updatestatus}
  \State $timer = timer + 1$; \label{l:timer} \label{l:inc}
  \State $p = taskList$;
  \While{$p$} \label{l:startrun1st}
    \If{$p\rightarrow status == \textit{INTERRUPT}$}
      \State \Return; \label{l:return}
    \ElsIf{$p\rightarrow status == \textit{READY}$}      
      \State $p\rightarrow status = \textit{RUNNING}$;
      \State \Call{$int\_on$}{$ $}; \Comment{to enable interrupts} \label{l:endrun1st}
      \State $p\rightarrow function()$; \Comment{to execute the task} \label{l:function}
      \State \Call{$int\_o\!f\!\!f$}{$ $};
      \State $p\rightarrow status = \textit{DORMANT}$;
    \EndIf
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
\Function{$updateStatus$}{$p$}
  \While{$p$}
    \If{$p\rightarrow status == \textit{RUNNING}$} \label{l:startupdate}
      \State $p\rightarrow status = \textit{INTERRUPT}$;
    \EndIf
    \If{$timer~\%~(p\rightarrow period) == 0$} \Comment{the task should be initiated}
      \If{$p\rightarrow status == \textit{DORMANT}$} \Comment{the previous job finishes} 
        \State $p\rightarrow status = \textit{READY}$;
      \Else \Comment{the status is \textit{READY} or \textit{INTERRUPT}}
	\State \Call{$reportTaskError$}{$p$}; \Comment{the task misses its deadline}
      \EndIf
    \EndIf \label{l:endupdate}
    \State $p = p\rightarrow next$;
  \EndWhile
\EndFunction
  \end{algorithmic}
\hrule
  \caption{The C-Like Pseudocode of $schedule()$}
  \label{a:schedule}
\end{figure}

In Figure~\ref{a:schedule}, the handler function $scheduler()$ first
updates status of all tasks in $taskList$ via function
$updateStatus()$. This updating actually initiates tasks which should
be scheduled in the current interrupt cycle. Then $scheduler()$
traverses the list to execute the ready tasks one by one, or to do a
return when encountering an interrupted\footnote{Note that the status
  \textit{INTERRUPT} indicates the task is interrupted for the moment,
  or was interrupted before but its execution is not complete yet.}
task. As for the function $updateStatus()$, it updates each task in
two steps: firstly, if the task is running, it becomes interrupted;
secondly for the task at its initiation time, if its previous job is
complete, it would be set to be ready, otherwise it misses its
deadline, producing an error. Notice that $schedule()$ is invoked only
when the interrupt request is handled, not when the interrupt is
disabled (the interrupt masking bit is set). Due to the interrupt
masking bit, its execution cannot be interrupted when it is updating
status of tasks or finding the next task to execute, however, it can
be interrupted while executing some task (Line~\ref{l:function}). This
allows the execution of $schedule()$ to be nested.

For simplicity, in the rest of this paper, we use \emph{scheduling} to
refer to the stage, from the moment when a pending interrupt request
is detected, to the moment when the first should-be-run periodic task
starts executing, i.e. Line~\ref{l:return} or~\ref{l:function} in
Figure~\ref{a:schedule}. Therefore, \emph{scheduling time} mainly
consists of three parts: (i) the time for switching context from the
running task, possibly none, to $schedule()$ when an interrupt request
is handled, (ii) the time spent by $schedule()$ finding and setting
the \emph{first} should-be-run periodic task
(Lines~\ref{l:1stline}-\ref{l:endrun1st} in Figure~\ref{a:schedule}),
and (iii) the time for switching context from $schedule()$ to that
task. \emph{Switching} refers to the stage, from the moment when a
periodic task completes its execution, to the moment when the next
should-be-run periodic task starts executing. \emph{Switching time}
thus also consists of three parts: (i) the time for switching context
from the complete task back to $schedule()$, (ii) the time spent by
$schedule()$ finding and setting the \emph{next} should-be-run
periodic task, and (iii) the time for switching context from
$schedule()$ to that task.

\section{Formal Modeling of the Implementation}
\label{s:formalism}
Considering some technical details of the platform, such as interrupt
masking mechanism, we model the implementation under the following
assumptions:

(A1') Jobs corresponding to task $\tau_i$ are initiated at the
beginning of scheduling that handles the requests triggered at times
$kT_i$ with integers $k\ge 0$.

(A2') Computation requirement $C_i$ for each task $\tau_i$ is not
constant. It varies randomly between minimum $C^{min}_i$ and maximum
$C^{max}_i$, i.e. within an interval $[C^{min}_i, C^{max}_i]$.

(A3) Tasks are independent, such that they are ready to run at their
initiation times and can be preempted instantly.

(A4') Scheduling time and switching time are considered, while other
overhead is ignored.

These assumptions make our model different from the standard one. For
instance, with assumption~(A1'), an interrupt request that occurs
while switching is performed will be pending, such that jobs of task
$\tau_i$ cannot initiate at $kT_i$. They should wait until switching
finishes and the interrupt masking bit is cleared, which is different
from~(A1). Note that (A3) says that jobs are ready to run at their
initiation times, however, no one can start running at exactly its
initiation time, because scheduling takes time. Under these
assumptions, an example for showing the execution of tasks scheduled
by the implementation is depicted in Figure~\ref{sf:rmsimp}. Note that
the second job of $\tau_1$ is initiated at time $5.5$ instead of $5$,
because switching is performed and the interrupt masking bit is set at
time $5$.

\begin{figure}[!t]
\centering
\subfloat[RMS Algorithm]{
  \begin{tikzpicture}[font=\small,scale=1.25,
    % styles
    system task/.style={fill=gray!70,draw=gray},
    periodic task/.style={fill=gray!70!black,draw=gray!50!black},
    task line/.style={gray!30,very thin},
    time line/.style={gray!50,very thin,dashed},
    short line/.style={gray!50,very thin,dashed},
    init/.style={{Triangle[fill=white,scale=1.1]}-},
    complete/.style={{Triangle[scale=1.1]}-}]

  % legend
  \begin{scope}[xshift=-2cm]  
    \draw [fill=gray!25,draw=gray!50,dashed] (1.5,2.3) rectangle +(4,-1);
    \draw (1.5,2.3) node [below right] {LEGEND};
    \draw [init] (2,1.85) -- +(0,-0.01);
    \draw (2.1,2) node [below right] {: Initiation of Tasks};
    \draw [complete] (2,1.42) -- +(0,0.1);
    \draw (2.1,1.7) node [below right] {: Completion of Tasks};
  \end{scope}

    % grid
    \foreach \x in {0.3,0.7}
      \draw[task line] (0,\x) -- (5.2,\x);

    % axes
    \draw[->,thick] (0,0) node [below] {$0$} -- (5.3,0) node [above] {time};
    \draw[thin] (0,0) -- (0,1);

    % time indicator
    \draw [time line] (1.5,0.7) -- (1.5,0);
    \draw (1.5,0) node [below] {$3$} -- +(0,2pt);
    \draw [time line] (2.5,0.7) -- (2.5,0);
    \draw (2.5,0) node [below] {$5$} -- +(0,2pt);
    \draw [time line] (4,0.7) -- (4,0);
    \draw (4,0) node [below] {$8$} -- +(0,2pt);
    \draw [time line] (4.5,0.3) -- (4.5,0);
    \draw (4.5,0) node [below] {$9$} -- +(0,2pt);
    \draw [time line] (5,0.7) -- (5,0);
    \draw (5,0) node [below] {$10$} -- +(0,2pt);

    % tasks
    \draw [periodic task] (0,0.7) rectangle +(1.5,0.14);
    \draw [periodic task] (1.5,0.3) rectangle +(1,0.14);
    \draw [periodic task] (2.5,0.7) rectangle +(1.5,0.14);
    \draw [periodic task] (4,0.3) rectangle +(0.5,0.14);

    % text
    \draw (0,0.75) node [left] {$\tau_1$};
    \draw (0,0.35) node [left] {$\tau_2$};
    \path (0,0.75) node [left,text opacity=0] {scheduling};
    \path (0,0.35) node [left,text opacity=0] {switching};

    % initiation of tasks
    \foreach \x in {0,2.5,5}
      \draw [init] (\x,0.7) -- +(0,-0.01);
    \foreach \x in {0,5}
      \draw [init] (\x,0.3) -- +(0,-0.01);

    % completion of tasks
    \draw [complete] (1.5,0.84) -- +(0,0.01);
    \draw [complete] (4,0.84) -- +(0,0.01);
    \draw [complete] (4.5,0.44) -- +(0,0.01);
  \end{tikzpicture}
\label{sf:rmsalg}}\\
\subfloat[RMS Implementation]{
  \begin{tikzpicture}[font=\small,scale=1.25,
    % styles
    system task/.style={fill=gray!70,draw=gray},
    periodic task/.style={fill=gray!70!black,draw=gray!50!black},
    task line/.style={gray!30,very thin},
    time line/.style={gray!50,very thin,dashed},
    short line/.style={gray!50,very thin,dashed},
    init/.style={{Triangle[fill=white,scale=1.1]}-},
    complete/.style={{Triangle[scale=1.1]}-}]

    % grid
    \foreach \x in {0.3,0.7,1.1,1.5}
      \draw[task line] (0,\x) -- (5.2,\x);

    % axes
    \draw[->,thick] (0,0) node [below] {$0$} -- (5.3,0) node [above] {time};
    \draw[thin] (0,0) -- (0,1.8);

    % time indicator
    \draw [time line] (0.5,1.5) -- (0.5,0);
    \draw (0.5,0) node [below] {$1$} -- +(0,2pt);
    \draw [time line] (1.25,1.5) -- (1.25,0);
    \draw (1.25,0) node [below] {$2.5$} -- +(0,2pt);
    \draw [time line] (1.75,1.1) -- (1.75,0);
    \draw (1.75,0) node [below] {$3.5$} -- +(0,2pt);
    \draw [time line] (2.25,1.1) -- (2.25,0);
    \draw (2.25,0) node [below] {$4.5$} -- +(0,2pt);
    \draw [time line] (2.75,1.5) -- (2.75,0);
    \draw (2.75,0) node [below] {$5.5$} -- +(0,2pt);
    \draw [time line] (3.25,1.5) -- (3.25,0);
    \draw (3.25,0) node [below] {$6.5$} -- +(0,2pt);
    \draw [time line] (4.25,1.5) -- (4.25,0);
    \draw (4.25,0) node [below] {$8.5$} -- +(0,2pt);
    \draw [time line] (4.75,0.3) -- (4.75,0);
    \draw (4.75,0) -- +(0,2pt);
    \draw (4.65,0) node [below] {$9.5$};
    \draw [time line] (5,1.5) -- (5,0);
    \draw (5,0) node [below] {$10$} -- +(0,2pt);

    % tasks
    \draw [system task] (0,0.7) rectangle +(0.5,0.14);
    \draw [periodic task] (0.5,1.5) rectangle +(0.75,0.14);
    \draw [system task] (1.25,0.3) rectangle +(0.5,0.14);
    \draw [periodic task] (1.75,1.1) rectangle +(0.5,0.14);
    \draw [system task] (2.25,0.3) rectangle +(0.5,0.14);
    \draw [system task] (2.75,0.7) rectangle +(0.5,0.14);
    \draw [periodic task] (3.25,1.5) rectangle +(1,0.14);
    \draw [system task] (4.25,0.3) rectangle +(0.5,0.14);

    % text
    \draw (0,1.55) node [left] {$\tau_1$};
    \draw (0,1.15) node [left] {$\tau_2$};
    \draw (0,0.75) node [left] {scheduling};
    \draw (0,0.35) node [left] {switching};

    % initiation of tasks
    \foreach \x in {0,2.75,5}
      \draw [init] (\x,1.5) -- +(0,-0.01);
    \foreach \x in {0,5}
      \draw [init] (\x,1.1) -- +(0,-0.01);

    % completion of tasks
    \draw [complete] (1.25,1.64) -- +(0,0.01);
    \draw [complete] (2.25,1.24) -- +(0,0.01);
    \draw [complete] (4.25,1.64) -- +(0,0.01);
  \end{tikzpicture}
\label{sf:rmsimp}}
\caption{A task set scheduled under RMS algorithm and the
  implementation, respectively. $T=T_1=5$, $T_2=10$.  In
  Figure~\ref{sf:rmsalg}, with assumption~(A2), $C_1=3$, $C_2=3$. In
  Figure~\ref{sf:rmsimp}, with assumption~(A2'), $C_1\in [1,3]$ and
  $C_2\in [1,3]$, assuming both scheduling time and switching time are
  $1$.}
\label{f:example}
\end{figure}
